Windowed least-squares model reduction for dynamical systems

### 2019-10-07 section4/numerical_methods.tex
thru p 15

\section{Numerical solution techniques}\label{sec:numerical_techniques}
The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
As \methodAcronym\ displays commonalities with optimal control problems, a large existing body of work can be leveraged. This section discuses a variety of such techniques and outlines how they can be used to solve the \methodAcronym\ system. We focus primarily on direct and indirect methods for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%Equation~\ref{eq:obj_gen_slab} displays commonalities with optimal control problems. For timeslabs $n = 1,\ldots,\nslabs$, the objective is to find the \textit{function} $\genstate^n(t)$ that minimizes the \textit{functional} $ \mathcal{J}^n$. The distinction between Eq.~\ref{eq:obj_gen_slab} and traditional optimal control problems is that, in the form presented in Eq.~\ref{eq:obj_gen_slab}, (1) there is not a controller and (2) there is no system that describes the dynamics for the generalized coordinates. 
%Techniques used to solve optimal control problems can be leveraged to solve Eq.~\ref{eq:obj_gen_slab} (or, equivalently, Eqs.~\eqref{eq:tclspg_oc1} and~\eqref{eq:tclspg_oc2}).

\subsection{\spatialAcronym: direct and indirect methods}
Numerical techniques to
solve optimal control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Direct method eschew the calculus of variations 
and attempt to ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus ``discretize then optimize."
Indirect methods, on the other hand, leverage the calculus of variations to
derive the Euler--Lagrange equations, which comprise the first-order optimality conditions~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs} associated with the objective 
functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
solve the optimization problem
``indirectly". For both direct and indirect methods, a variety of
discretization/solution approaches are then possible. Collocation methods,
finite elements, spectral methods, shooting methods, etc., are all examples of
plausible solution techniques.

In the present context, both direct and indirect methods are investigated to solve \methodAcronym\ with \spatialAcronym\ trial subspaces. We consider:

\begin{itemize} \item Discretize then Optimize (Direct Methods): 
\begin{itemize}
\item Discretize and then optimize approaches are outlined for linear multistep methods. 
\end{itemize} 
\item Optimize then Discretize (Indirect Methods): 
\begin{itemize} 
\item An indirect method leveraging the forward-backward sweep algorithm is outlined 
to solve~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}. 
 \end{itemize} 
\end{itemize} 
It
is emphasized that a variety of other approaches exist. A detailed exploration
of additional approaches will be a subject of future work. Section~\ref{sec:direct} 
will outline direct solution approaches based on linear multistep methods, 
while Section~\ref{sec:indirect} will outline the indirect solution approach.

>>>>

\section{Numerical solution techniques}\label{sec:numerical_techniques}
The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
\KTC{Passive voice:} As \methodAcronym\ displays commonalities with optimal
control problems, a large existing body of work can be leveraged. This section
discuses a variety of such techniques and outlines how they can be used to
solve the \methodAcronym\ system. We focus primarily on \KTC{earlier we kept
saying 
direct (i.e.,
	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
	Let's be consistent when we re-introduce it here.}
optimize direct and indirect methods for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%Equation~\ref{eq:obj_gen_slab} displays commonalities with optimal control problems. For timeslabs $n = 1,\ldots,\nslabs$, the objective is to find the \textit{function} $\genstate^n(t)$ that minimizes the \textit{functional} $ \mathcal{J}^n$. The distinction between Eq.~\ref{eq:obj_gen_slab} and traditional optimal control problems is that, in the form presented in Eq.~\ref{eq:obj_gen_slab}, (1) there is not a controller and (2) there is no system that describes the dynamics for the generalized coordinates. 
%Techniques used to solve optimal control problems can be leveraged to solve Eq.~\ref{eq:obj_gen_slab} (or, equivalently, Eqs.~\eqref{eq:tclspg_oc1} and~\eqref{eq:tclspg_oc2}).

\subsection{\spatialAcronym: direct and indirect methods}
Numerical techniques to
solve optimal control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}.  Direct \KTC{Typo:} method eschew
the calculus of variations \KTC{weird to say they `eschew' calculus. Find
something better here}
and attempt to ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus ``discretize then optimize."
Indirect methods, on the other hand, leverage the calculus of variations to
derive the Euler--Lagrange equations, which comprise the first-order
optimality conditions\KTC{use endash}~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs} associated with the objective 
functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
solve the optimization problem
``indirectly". \KTC{Passive voice:} For both direct and indirect methods, a variety of
discretization/solution approaches are then possible. Collocation methods,
finite elements, spectral methods, shooting methods, etc., are all examples of
plausible solution techniques.

In the present context, both direct and indirect methods \KTC{Passive voice:}
are investigated to solve \methodAcronym\ with \spatialAcronym\ trial
subspaces. We consider: \KTC{Weird capitalization and bulleting. Make
consistent with earlier in the paper, e.g., introduction, and don't break into
sub-bullet, looks ugly.}

\begin{itemize} \item Discretize then Optimize (Direct Methods): 
\begin{itemize}
\item Discretize and then optimize approaches are outlined for linear multistep methods. 
\end{itemize} 
\item Optimize then Discretize (Indirect Methods): 
\begin{itemize} 
\item An indirect method leveraging the forward-backward sweep algorithm is outlined 
to solve~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}. 
 \end{itemize} 
\end{itemize} 
\KTC{Passive voice:} It
is emphasized that a variety of other approaches exist. A detailed exploration
of additional approaches is the subject of future work.
Section~\ref{sec:direct}  outlines direct solution approaches based on linear
multistep methods, while Section~\ref{sec:indirect} outlines the indirect
solution approach.

### 2019-10-07 section4/spatial_numerical.tex
thru p 15

\subsection{\spatialAcronym: direct solution approach}\label{sec:direct} 
Direct approaches solve~\eqref{eq:tclsrm} by
``transcribing" the infinite dimensional optimization problem into a finite
dimensional one by discretizing the state and objective function in time.
The minimization problem is then reformulated as a nonlinear programming, or
optimization, problem. A variety of direct solution approaches exist, 
including collocation approaches, orthogonal polynomial (spectral  method) 
approaches, and genetic algorithms.

In the present context, the most straightforward direct solution approach consists of the 
following steps:
\begin{enumerate}
\item Numerically discretize the FOM ODE (and hence the \textit{integrand} of~\eqref{eq:obj_gen_slab}).
\item Select a numerical quadrature rule to evaluate the \textit{integral}
in~\eqref{eq:obj_gen_slab}.
\end{enumerate}
To achieve this agenda, first define a discrete temporal grid 
by dividing each time window $[\timeStartArg{n},\timeEndArg{n}]$ into
$\nstepsArg{n}$ strictly increasing time-step instances described by
$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$ with  
\begin{align}
&\timeWindowArg{n}{0} \le \ldots \le \timeWindowArg{n}{\nstepsArg{n}} , \label{eq:timegrid1} \;
\timeWindowArg{n}{0} \defeq \timeStartArg{n},\;
\timeWindowArg{n}{\nstepsArg{n}} \defeq \timeEndArg{n} . 
%\timeWindowArg{n}{0} = \timeStartArg{n}, \timeWindowArg{n}{\nstepsArg{n}} =
%\timeEndArg{n}
\end{align}
This discretization is depicted in Figure~\ref{fig:slab_fig2}.
We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
other time-stepping methods (e.g., Runge--Kutta) follows closely. 
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
\caption{Decomposition of time windows into time-step instances.} 
\label{fig:slab_fig2} 
\end{centering} 
\end{figure}
%We now outline several commonly employed temporal discretization 
%schemes, specifically linear multistep and collocation (Runge--Kutta) methods, that leverage this time-grid to discretize the FOM ODE and 
%objective functional.

@@

>>>>

\subsection{\spatialAcronym: direct solution approach}\label{sec:direct} 
\KTC{Does the time grid definted in this section have any relationship to the
time grid used for time discretization? Haven't gotten past HERE below, but
wanted to check (maybe it's made clear later).}
Direct approaches solve~\eqref{eq:tclsrm} by
``transcribing" the infinite dimensional optimization problem into a finite
dimensional one by discretizing the state and objective function in time.
\KTC{Passive voice:} The minimization problem is then reformulated as a nonlinear programming, or
optimization, problem. A variety of direct solution approaches exist, 
including collocation approaches, orthogonal polynomial \KTC{is this a
spectral method, or the spectral method? clarify} (spectral  method) 
approaches, and genetic algorithms.

In the present context, the most straightforward direct solution approach consists of the 
following steps: \KTC{Keep this inline. Don't see a benefit to multi-lining a
two-step algorithm}
\begin{enumerate}
\item Numerically discretize the FOM ODE (and hence the \textit{integrand} of~\eqref{eq:obj_gen_slab}).
\item Select a numerical quadrature rule to evaluate the \textit{integral}
in~\eqref{eq:obj_gen_slab}.
\end{enumerate}
To achieve this agenda \KTC{agenda? odd way to say it.}, first define a discrete temporal grid 
by \KTC{you can't devide (i.e., decompose) a window into instances. you can divide a window into
sub-windows. you can also simply define a set of time instances in the window.
no need to emphasize `strictly increasing', as any set of points in the window
will by necessity be ordered.} dividing each time window $[\timeStartArg{n},\timeEndArg{n}]$ into
$\nstepsArg{n}$ strictly increasing time-step instances described by
$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}$ with  \KTC{use cdots, not
hdots or ldots. Don't use $\defeq$ for the first and last times; these are
just equalities, i.e., you happen to set them equal to those values, but they
aren't identically the same:}
\begin{align}
&\timeWindowArg{n}{0} \le \ldots \le \timeWindowArg{n}{\nstepsArg{n}} , \label{eq:timegrid1} \;
\timeWindowArg{n}{0} \defeq \timeStartArg{n},\;
\timeWindowArg{n}{\nstepsArg{n}} \defeq \timeEndArg{n} . 
%\timeWindowArg{n}{0} = \timeStartArg{n}, \timeWindowArg{n}{\nstepsArg{n}} =
%\timeEndArg{n}
\end{align}
\KTC{Passive voice:} This discretization is depicted in Figure~\ref{fig:slab_fig2}.
We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
other time-stepping methods (e.g., Runge--Kutta) follows closely. 
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
	\caption{Decomposition of time windows into time-step instances. \KTC{You
	can't decompose a time interval into time instances. You can define time
	instances on that interval.} \KTC{Don't call them `time-step instances'.
	This woudl be an instance of a time step. These are `time instances',
	correct? If so, fix this terminology everywhere.}} 
\label{fig:slab_fig2} 
\end{centering} 
\end{figure}
%We now outline several commonly employed temporal discretization 
%schemes, specifically linear multistep and collocation (Runge--Kutta) methods, that leverage this time-grid to discretize the FOM ODE and 
%objective functional.

@@

\KTC{HERE}
